Learning by Teaching a Robot: the Case of Handwriting

### 2015-07-10 main.tex
diego re-writing
Proof-reading first half

    Robots for education are not limited to support ICT teaching, and robots are
    finding new ways to enter and blend in the classroom. This article reports
    on such a new paradigm for educative robots, that involves \emph{learning by
    teaching} and strong \emph{social engagement} to help children struggling
    with handwriting. Our system relies on machine-learning and child-robot
    interaction with a small humanoid robot, and we present several real-world
    studies in schools and with occupational therapists that led us to promising
    initial results.

@@

While Henry is lively and always quick at shifting his attention from one
activity to another, Diego is shy and poised. Two very different children,
facing however the same difficulty to write in a legible manner. And, hidden
beyond this impaired skill, psycho-social difficulties also arise: they
underperform at school, Henry has to go for follow-up visits every week,...
they live under the label ``requires special care''. This is a source of
anxiety, for the children, and for their parents alike.

@@

We present in this article a new take on this educative challenge, as a
remediation procedure that involve a \emph{bad writer} robot that is taught by the
child: by building on the \emph{learning by teaching} paradigm, not only the
child practises hand-writing but, as (s)he takes on the role of the teacher,
(s)he also positively reinforce his/her self-esteem and motivation: his/her
social role shifts from the ``underperformer'' to ``the one who knows and teaches''.

@@

The \emph{learning by teaching} paradigm, which engages the student in the act
of teaching another, has been shown to produce motivational, meta-cognitive, and
educational benefits in a range of disciplines~\cite{Rohrbeck2003}.  The
application of this paradigm to handwriting intervention remains, however,
unexplored. One reason for this may be due to the requirement of an
appropriately unskilled peer for the child to tutor: it may prove difficult if
the child is the lowest performer in his/her class.  In some cases, it may be
appropriate for a peer or teacher to simulate a na\"ive learner for the target
child to teach. For handwriting however, where one's skill level is visually
evident, this acting is likely to be rapidly detected. This motivates the use of
an artificial teachable agent which can be configured for a variety of skill
levels, and for which children do not have preconceptions about its handwriting
ability.

@@

The work presented here investigates a particular role for a robot in the
education of handwriting: not only is the robot actively performing the activity
by drawing letters, but it does so in a way that engages the child in a very
specific social role. The child is the teacher in this relationship and the
robot is the learner: the child must engage in a (meta-) cognitive relationship
with the robot to try to understand why the robot fails and how to help it best.
Here, the robot is more than just an activity facilitator or orchestrator -- its
physical presence and embodiment induce agency and anthropomorphising, and
cognitively engage the child into the learning activity.

Besides that, the commitment of the child into the interaction build on the
``protégé effect'': teachable computer-based agents have previously been used to
encourage the ``protégé effect'', wherein students invest more effort into
learning when it is for a teachable agent than for themselves~\cite{Chase2009}.
As we are concerned with learning of a physical skill, the learning agent
developed is embodied in a humanoid robot which is capable of physically
demonstrating handwriting trajectories to its child learning partner.  This is
supported by the potential for motor mimicry to yield significant improvements
in handwriting interventions in which letter formations are demonstrated to
participants~\cite{Berninger1997}. Furthermore, when compared to screen-based
agents, robotic partners have been shown in some contexts to increase users'
compliance with tasks~\cite{Bainbridge2011}, maintain more effective long-term
relationships~\cite{Kidd2008}, and produce greater learning gains when acting as
tutors~\cite{Leyzberg2012}.

@@

In our case, the role of agency is stronger: it induces meta-cognition (``I am
interacting with an agent, so I need to reflect on how to best teach him''),
which is beneficial for the learning process; it also induces a protégé effect
(``I want my robot-agent to succeed!'') which support the commitment of child
into the interaction, also for longer period of time (we report in this article
upon to case studies where children were interacting with the same robot for several
hours, spanned over several weeks).

\subsection*{Overview of the article}

We have sofar frame our intents: to design a robotic system that would support
handwriting remediation in an original way: by getting children to teach a
robotic agent how to write, those children would both practise without knowing
it and recover self-confidence and self-esteem by supporting a
worst-than-themselves robotic ``student''.

The following sections of the paper go into details. The next section provides a
brief overview of the robotic system and the interaction it induces. We then
give the details of machine-learning techniques that allow the robot to learn
from the children, followed by the implementation on the Nao robot.

Next, we present and report on the four school experiments and the two case
studies that we conducted over the last two years. While the focus of the school
experiments was mostly technical validation and data acquisition, the two case
studies that lasted each about 4 weeks, gave us initial insights on the
relevance and effectiveness of our approach.

@@

Figure~\ref{experimental_setup} illustrates our general experimental setup: a
face-to-face child-robot interaction with an (autonomous) Aldebran's Nao robot.

A tactile tablet (with a custom application) is used for both the robot and the
child to write: during a typical round, the child requests the robot to write
something (a single letter, a number or a full word), and push the tablet
towards the robot, the robot writes on the tablet by gesturing the writing, but
\emph{without actually physically touching the tablet}), the child then pull
back the tablet, correct the robot's attempt by writing him/herself on top or
next to the robot's writing (see the picture in Figure~\ref{fig:diego}), and
``send'' his/her demonstration to the robot by pressing a small button on the
tablet. The robot ``learns'' from this demonstration and tries again.

@@

Implementing such a system raises several challenges: first, acquiring,
analyzing and learning from hand-written demonstration lays at the core of the
our approach and we developed several algorithms for the robot to generate
initial bad writing and to respond in an adequate manner, showing visible (but
not too quick) writing improvements.

Then, the actual implementation on the robot required the coordination of
several modules (from performing gestures and acquiring the user's input to
the high-level state machine), spread over several devices (the robot itself,
one laptop and up to four tactile tablets for certain studies we conducted). We
relied on ROS to ensure the synchronization and communication between these
modules.

@@

The main insight for both the generation and the learning of letters is to
reason in the eigenspace of the letters, instead of the cartesian space. The
eigenspace of each letters is spanned by the first $n$ eigenvectors (in our
experiments, $3 < n < 6$) of the covariance matrix observed in a standard
dataset of adult letters (the UJI Pen Characters 2 dataset~\cite{Llorens2008}).
This procedure, based on a Principle Component Analysis (PCA), is explained in
details in a previously published article~\cite{hood2015when}.

By changing the eigenvalues associated to these eigenvectors and regenerating
letters with the reverse procedure, it becomes then easy to generate new
letters, with distortions that are actually plausible handwriting errors: they
are actually \emph{exaggerations} of variance of writing styles that naturally
exist amongst adult writers.  Figure~\ref{fig:sampleLetters} shows examples of
deformed ``g'' generated with such a technique.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{cowriter-g}
    \caption{\small \label{fig:sampleLetters} \textbf{Generating bad letters}:
        effect of varying the first five eigenvalues (rows) of the shape model
        of ``g'' by different factors (columns). The percentage of the total
        variance in the dataset explained by each eigenvalue is shown      on
        the appropriate row. Examples noted A, B, C, D illustrate how the
        PCA-based approach allows to \emph{automatically} generate letters whose
        errors can be \emph{semantically} interpreted: A has a too large bottom
        loop, B has a wide top loop, the bottom loop of C is not correctly
        closed, the top loop of D is not closed.}

@@

The same technique can be applied to \emph{classify} demonstrations,
\emph{assess} their quality and
\emph{learn} from them. Figure~\ref{fig:h} shows 9 allographs of ``h'' written by a 6
years old child, along with the mean letter from the
dataset (reference letter). By projecting each of the demonstrations onto the eigenspace
of ``h'' (Figure~\ref{fig:h}b), we observe that:

@@

    \caption{\small Projecting demonstrated letters onto the eigenspace
    generated from the reference dataset effectively clusters the samples
    according to their topological similarity. Allographs that are visually similar to
    the reference are close to it in the eigenspace.}
    \label{fig:h}
\end{figure}

The algorithm for machine-learning becomes then a simple matter of converging at
a specific pace towards the child demonstration in the eigenspace.
Figure~\ref{learning_6_demos}, at the end of the article, shows a complete learning cycle
with the number ``6''.

@@

Our system is embodied through an Aldebaran's {\sc nao} (V4 or V5, depending on the studies)
humanoid robot. This choice is motivated by its approachable
design~\cite{Gouaillier2008}, its size (58cm) and inherently safe structure (lightweight
plastic) (making it suitable for close interaction with children), its low price
(making it closer to what school may afford in the coming years) and finally its
ease of deployment on the field.

{\sc nao} is a humanoid biped, has 25 degrees of freedom,
two cameras, speech capabilities and the ability to autonomously execute a range
of tasks. While it can walk, this capability is not used in this project, and
the robot remains in a crouching posture during the whole interaction.

Robotic handwriting requires precise closed-loop control of the arm and hand
motion. Because of the limited fine motor skills possible
with such an affordable robot, in addition to the absence of force feedback and
other technical necessities, we have opted for \emph{simulated
handwriting}: the robot draws letters in the air, and the actual writing is
displayed on a synchronised tablet.

>>>>

    Robots for education are not limited to support ICT teaching, and they are
    indeed finding new roles in the classroom. This article reports on such a
    new paradigm for educative robots, that involves \emph{learning by teaching}
    and strong \emph{social engagement} to help children struggling with
    handwriting. Our system relies on machine-learning and child-robot
    interaction with a small humanoid robot, and we present several real-world
    studies in schools and with occupational therapists that led us to promising
    initial results.

@@

While Henry is lively and always quick at shifting his attention from one
activity to another, Diego is shy and poised. Two very different children,
facing however the same difficulty to write in a legible manner. And, hidden
beyond this impaired skill, psycho-social difficulties also arise: they
underperform at school, Henry has to go for follow-up visits every week,
they both live under the label ``requires special care''. This is a source of
anxiety for the children and for their parents alike.

@@

We present in this article a new take on this educative challenge, as a
remediation procedure that involve a ``bad writer'' robot that is \emph{taught}
by the child: by building on the \emph{learning by teaching} paradigm, not only
the child practises hand-writing but, as (s)he takes on the role of the teacher,
(s)he also positively reinforce his/her self-esteem and motivation: his/her
social role shifts from the ``underperformer'' to ``the one who knows and
teaches''.

@@

The \emph{learning by teaching} paradigm, which engages the student in the act
of teaching another, has been shown to produce motivational, meta-cognitive, and
educational benefits in a range of disciplines~\cite{Rohrbeck2003}.  The
application of this paradigm to handwriting intervention remains, however,
unexplored. One reason for this may be due to the requirement of an
appropriately unskilled peer for the child to tutor: this may indeed prove
difficult if the child is the lowest performer in the class.  In some cases, it
may be appropriate for a peer or teacher to simulate a na\"ive learner for the
child to teach. For handwriting however, where one's skill level is visually
evident, this acting is likely to be rapidly detected. This motivates the use of
an artificial teachable agent which can be configured for a variety of skill
levels, and for which children do not have preconceptions about its handwriting
ability.

@@

We also investigate here a particular role for a robot in the
education of handwriting: not only is the robot actively performing the activity
by drawing letters, but it does so in a way that engages the child in a very
specific social role. The child is the teacher in this relationship and the
robot is the learner: the child must engage in a (meta-) cognitive relationship
with the robot to try to understand why the robot fails and how to help it best.
Here, the robot is more than just an activity facilitator or orchestrator -- its
physical presence and embodiment induce agency and anthropomorphising, and
cognitively engage the child into the learning activity.

Besides the commitment of the child into the interaction build on the
``prot\'eg\'e effect'': the teacher feels responsible for his student, commits
to the student's success and possibly experiences student's failure as his own
failure to teach. Teachable computer-based agents have previously been used to
encourage this ``prot\'eg\'e effect'', wherein students invest more effort into
learning when it is for a teachable agent than for themselves~\cite{Chase2009}.
We rely on this cognitive mechanism to reinforce the child's commitment into the
robot-mediated handwriting activity.

@@

In our case, the role of agency is stronger: it induces meta-cognition (``I am
interacting with an agent, so I need to reflect on how to best teach him'')
which is beneficial for the learning process; it also induces a prot\'eg\'e
effect (``I want my robot-agent to succeed!'') which support the commitment of
child into the interaction, also for longer period of time (we report in this
article upon two case studies where children were indeed interacting with the
same robot for several hours, spanned over several weeks).

Building on these socio-cognitive mechanisms, our intent is therefore to design
a robotic system that would effectively support handwriting remediation in an
original way: by getting children to teach a robotic agent how to write, those
children would both practise without knowing it and recover self-confidence and
self-esteem by supporting a worst-than-themselves robotic ``student''.

The following sections of the paper go into details. We first provides a
brief overview of the robotic system and the interaction it induces. We then
give the details of machine-learning techniques that allow the robot to learn
from the children, followed by the implementation on a {\sc nao} robot.

We then present and report on the four school experiments and the two case
studies that we conducted over the last two years. While the focus of the school
experiments was mostly technical validation and data acquisition, the two case
studies that lasted each about 4 weeks, gave us initial insights on the
relevance and effectiveness of our approach.

@@

Figure~\ref{experimental_setup} illustrates our general experimental setup: a
face-to-face child-robot interaction with an (autonomous) Aldebran's {\sc nao}
robot.

A tactile tablet (with a custom application) is used for both the robot and the
child to write: during a typical round, the child requests the robot to write
something (a single letter, a number or a full word), and push the tablet
towards the robot, the robot writes on the tablet by gesturing the writing (but
without actually physically touching the tablet), the child then pull back the
tablet, corrects the robot's attempt by writing him/herself on top or next to
the robot's writing (see Figure~\ref{fig:diego}), and ``send'' his/her
demonstration to the robot by pressing a small button on the tablet. The robot
``learns'' from this demonstration and tries again.

@@

Implementing such a system raises several challenges: first, the acquisition,
analyse and learning from hand-written demonstration, which lays at the core of the
our approach, necessitates the development of several algorithms for the robot to generate
initial bad writing and to respond in an adequate manner, showing visible (but
not too quick) writing improvements.

Then, the actual implementation on the robot requires the coordination of
several modules (from performing gestures and acquiring the user's input to
the high-level state machine), spread over several devices (the robot itself,
one laptop and up to four tactile tablets for certain studies we conducted). We
relied on ROS to ensure the synchronization and communication between these
modules.

@@

The main insight for both the generation and the learning of letters is to
reason about the shape of letters in their eigenspace, instead of the natural
cartesian space. The eigenspace of each letters is spanned by the first $n$
eigenvectors (in our experiments, $3 < n < 6$) of the covariance matrix
generated from a standard dataset of adult letters (the UJI Pen Characters 2
dataset~\cite{Llorens2008}).  This procedure, based on a Principle Component
Analysis (PCA), is explained in details in a previously published
article~\cite{hood2015when}.

By changing the eigenvalues associated to these eigenvectors and regenerating
letters with the reverse procedure, it becomes then easy to generate new
letters, with distortions that are actually plausible handwriting errors: they
are actually \emph{exaggerations} of variance of writing styles that naturally
occurs amongst adult writers.  Figure~\ref{fig:sampleLetters} shows examples of
deformed ``g'' generated with such a technique.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{cowriter-g}
    \caption{\small \label{fig:sampleLetters} \textbf{Generating bad letters}:
        effect of varying the first five eigenvalues (rows) of the shape model
        of ``g'' by different factors (columns). The percentage of the total
        variance in the dataset explained by each eigenvalue is shown      on
        the appropriate row. Examples noted A, B, C, D illustrate how the
        PCA-based approach allows to \emph{automatically} generate letters whose
        errors can be \emph{semantically} interpreted, \eg A has a too large bottom
        loop, B has a wide top loop, the bottom loop of C is not correctly
        closed, the top loop of D is not closed, etc.}

@@

The same technique can also be applied to \emph{classify} demonstrations,
\emph{assess} their quality and \emph{learn} from them. Figure~\ref{fig:h} shows
for example 9 allographs of ``h'' written by a 6 years old child, along with the
mean letter from the dataset (reference letter). By projecting each of the
demonstrations onto the eigenspace of ``h'' (Figure~\ref{fig:h}b), we observe
that:

@@

    \caption{\small Projecting demonstrated letters onto the eigenspace
    generated from the reference dataset effectively clusters the samples
    according to their topological similarity. Allographs that are similar to
    the reference are close to it in the eigenspace.}
    \label{fig:h}
\end{figure}

The algorithm for machine-learning becomes then a simple matter of converging at
a specific pace towards the child demonstration in the eigenspace.
Figure~\ref{learning_6_demos}, at the end of the article, illustrates the
process with the complete learning cycle of the number ``6''.

@@

Our system is embodied in an Aldebaran's {\sc nao} (V4 or V5, depending on the
studies) humanoid robot. This choice is motivated by its approachable
design~\cite{Gouaillier2008}, its size (58cm) and inherently safe structure
(lightweight plastic) making it suitable for close interaction with children,
its low price (making it closer to what school may afford in the coming years)
and finally its ease of deployment on the field.

Robotic handwriting requires precise closed-loop control of the arm and hand
motion. Because of the limited fine motor skills possible with such an
affordable robot, in addition to the absence of force feedback, we have opted
for \emph{simulated handwriting}: the robot draws letters in the air, and the
actual writing is displayed on a synchronised tablet.
